---
title: "Critical Thinking - Module 1 - Option 2"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
 
author: "David Edwards"
date: "1/22/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Display the first 6 rows of the built-in Air Quality data set


```{r}
head(airquality)
```
```{r}
date()
```


## Display the summary statistics for all attributes (columns) of the Air Quality data set.

```{r airquality}
summary(airquality)
```

## Select any two attributes of the Air Quality data set. For each of your two selected attributes display:

### Ozone
#### Mean
```{r}
mean(airquality$Ozone, na.rm=TRUE)
```

#### Median
```{r}
median(airquality$Ozone, na.rm=TRUE)
```

#### Range
```{r}
range(airquality$Ozone, na.rm=TRUE)
```

#### Histogram
```{r}
Ozone <- airquality$Ozone[!is.na(airquality$Ozone)]
hist(Ozone, xlim=c(0,200), xlab="Ozone parts per billion")
boxplot(Ozone, main="Boxplot of Ozone", xlab="Ozone", ylab="Ozone parts per billion")
```

### Temp
#### Mean
```{r}
Temperature <- airquality$Temp[!is.na(airquality$Temp)]
mean(Temperature)
```

#### Median
```{r}
median(Temperature)
```

#### Range
```{r}
range(Temperature)
```

#### Histogram
```{r}
hist(Temperature, xlim=c(50,110), xlab="Temperature (\u00B0F)")
boxplot(Temperature, main="Temperature Boxplot", xlab="Temperature", ylab="Degrees Fahrenheit")
```

## For the same two attributes of the Air Quality data set that you selected in step 3, display the correlation between these two attributes
### statistically via a correlation number

Creating the corellation resulted in an NA because some Ozone levels were NA.  I filtered out the NA Ozone measurements, and the corresponding Temperature elements as well.
```{r}
Ozone <- airquality$Ozone[!is.na(airquality$Ozone)]
Temperature <- airquality$Temp[!is.na(airquality$Ozone)]
cor(Ozone, Temperature, method=c("pearson", "kendall", "spearman"))
```

### visually via a scatterplot diagram.

```{r}
plot(Ozone, Temperature, main="Scatterplot Temp and Ozone", xlab="Ozone (ppb) ", ylab="Temperature (\u00B0F)")
lines(lowess(Ozone, Temperature), col="blue")
```


